---
title: "Grade analysis"
author: "Ghina Mezher"
format: html
---

```{r message=FALSE}
here::i_am("r-101-grade-analysis.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
library(stringr)
```

## Question 1
```{r message=FALSE}
df<-vroom(here("grades.csv"))
na.rm=TRUE
```
## Question 2
```{r}
df|>
  summarise("Minimum"=min(`Exam`,na.rm=TRUE),"Maximum"=max(`Exam`,na.rm=TRUE),"Median"=median(`Exam`,na.rm=TRUE),"Mean"=mean(`Exam`,na.rm=TRUE))|>
  knitr::kable()
```
## Question 3
```{r}
sum(is.na(df$Exam))
```
`r sum(is.na(data$Exam)` students who did not pass the final exam is. 

## Question 4
```{r}
df|>ggplot(aes(x=Exam))+
  geom_histogram(binwidth=2,fill="darkred",color="red",na.rm=TRUE)+
  labs(title= "Distribution of Exam grades",
       x="Grades to the final exam",
      y="Frequency of grades")
```


## Question 5
```{r}
group_counts<-df|>
  group_by(Group)|>
  summarise(Number=n())

knitr::kable(group_counts)
```
## Question 6
```{r}
ggplot(group_counts,aes(x=Group,y=Number))+
  geom_col(fill="darkblue")+
  labs(title="Number of students per group",x="Groups",y="Number of students")
```
## Question 7
```{r}
#first option
ggplot(df,aes(x=Exam, color=Group))+
  geom_density()



#Second option
ggplot(df,aes(x=Group, y=Exam))+
  geom_boxplot(fill="darkgreen")+
  labs(title="Distribution of grades to the final exam according to groups",x="Groups",y="Grades to the final exam")+
  coord_flip()
```
The first graphical representation to be criticized.

## Question 8
```{r}
na_grades<-df|>
  group_by(Group)|>
  summarise("Number of students missing"=sum(is.na(Exam)))

knitr::kable(na_grades)
```
## Question 9
```{r}
ggplot(na_grades,aes(x=Group,y=`Number of students missing`))+
  geom_col(fill="purple")+
  labs(title="Missing students to the final exam by group", x="Groups",y="Number of missing students")+
  coord_flip()
```
## Question 10
```{r}
reshape_df<-df|>pivot_longer(cols=-c(Id,Group))
```

#Question 11
```{r}
NA_count<-reshape_df|>
  group_by(Id)|>
  summarise("Missing grade"=sum(is.na(value)))

knitr::kable(NA_count)
```
## Question 12
```{r}
NA_count|>
  ggplot(aes(x=Id,y=`Missing grade`))+
  geom_col()+
  labs(title="Missing grades per Student",
       x="Student Id",
       y="Missing grade count")
```

##Question 13
```{r}
missed_exam_group<-reshape_df|>
  filter(str_starts(name,"Exam"))|>
  group_by(Group)|>
  summarise("No grade for the exam"=sum(is.na(value)))
missed_exam_group
```


## Question 14
```{r}
NA_count_online<-reshape_df|>filter(str_starts(name,"Online_"))|>
  group_by(Id)|>
  summarise("Missing grade"=sum(is.na(value)))

knitr::kable(NA_count_online) 
```

## Question 15
```{r}
grades_NA<-df|>
  rowwise()|>
  mutate(NA_count=sum(is.na(c_across(starts_with("Online")))))
  
grades_NA_res<-grades_NA|>
  group_by(NA_count)|>
  summarise(mean=mean(Exam,na.rm=TRUE))

ggplot(grades_NA_res,aes(x=NA_count,y=mean))+
  geom_col()+
  ylim(0,20)+
  labs(title="Exam grades according to number of online tests missed",
       x="Numbers of online tests missed",
       y="Grade to the exam")
  
```
We add ylim(0,20) to take into account that the exam is graded over 20. However it makes reading the exact values of the exam variable for each column.

## Question 16
```{r}
missed<-df|>
  rowwise()|>
  mutate("Missed an online test"=any(is.na(c_across(starts_with("MCQ_")))))

missed_id<-missed|>
  select(Id,`Missed an online test`)

knitr::kable(missed_id)
  
```
## Question 17
```{r}
missed_group<-missed|>
  group_by(Group)|>
  summarise(P_missed=sum(`Missed an online test`)/n()*100)

knitr::kable(missed_group)
```
##Question 18
```{r}
Exam_group<-df|>
  group_by(Group)|>
  summarise(mean=mean(Exam,na.rm=TRUE))

Exam_missed<-inner_join(missed_group,Exam_group,by="Group")|>
  ggplot(aes(x=P_missed,y=mean,fill=Group))+
  geom_hex()+
  labs(title="Every group's average grade at the Exam according to the percentage of missed MCQ gradse",
       x="Percentage of students who missed at least one MCQ",
       y="Average grade to the final exam")
Exam_missed
```


